An Ant Cuticle Texture Classification Algorithm for Ecological Anaylsis

### 2021-09-03 paper/paper.tex
add overall class distribution

\maketitle

\section{Introduction}
Insects compose half of biodiversity and rank among the most dominant organisms
in terrestrial ecosystems \cite{sheikh_diverse_2017}. A key factor for the
ecological success of insects is their exoskeleton, also known as cuticle. The
cuticle protects insects from predation, provides structural support, prevents
desiccation, and serves as a canvas for advertising visual and chemical signals
\cite{gullan_insects_2009}. Research has heavily focused on the macrostructures
and internal chemical components that make the exoskeleton functional and more
recent work is being done to understand the functional aspects of external
cuticle micro sculpturing \cite{muthukrishnan_insect_2020,
  gunderson_insect_1989, watson_diversity_2017}.

@@

\begin{table}[h]
  \centering
  \caption{Dataset Class Distribution}
  \label{table:classdist}
  \include{tables/all_samples.tex}
\end{table}

Images were collected and classified by undergraduate students according to the
sculpturing identification protocol. Overall, there are a total of 2443 images
in the dataset. The class distribution is shown in Table \ref{table:classdist}.
The relatively small number of images in the dataset paired with the uneven
class distribution makes it difficult to perform a meaningful classification.
However, the goal of this project is to develop a classification algorithm that
can be used to help classify over 200,000 ant images. Therefore, it is desirable
to use a few representative images from each class to train the algorithm.
Additionally, data augmentation and other methods are performed to increase the
number of training images to create a more robust classification algorithm.

>>>>

\maketitle
\section{Introduction}
Insects compose half of biodiversity and rank among the most dominant organisms
in terrestrial ecosystems \cite{sheikh_diverse_2017}. A key factor for the
ecological success of insects is their exoskeleton, also known as cuticle. The
cuticle protects insects from predation, provides structural support, prevents
desiccation, and serves as a canvas for advertising visual and chemical signals
\cite{gullan_insects_2009}. Research has heavily focused on the macrostructures
and internal chemical components that make the exoskeleton functional and more
recent work is being done to understand the functional aspects of external
cuticle micro sculpturing \cite{muthukrishnan_insect_2020,
  gunderson_insect_1989, watson_diversity_2017}.

Texture is an important feature in many applications, such as image processing,
pattern recognition, and computer vision. Analysis of textures can be broken
into three main categories: texture classification, texture segmentation, and
texture synthesis. The process of classifying a texture into a set of categories
and relies on three different approaches: feature-based, model-based, or
structural \cite{reed_review_1993}. In this paper, we focus on a
\textit{model-based approach} which attempt to extract parameters to reveal
common patterns and use those parameters to automatically distinguish between
different textures \cite{maillard_texture_2003}.

Although there is some work regarding grouping ants into categories of similar
cuticle, automated classification has yet to become an active area of research.
Due to the large number of different ant species, the classification of ants
into categories of similar texture is difficult to accomplish manually. Texture
analysis has shown promising results in related fields, such as plant
identification \cite{boudra_plant_2018}. With modern texture analysis methods,
the classifications of ants can be automated and the results can be used to
study the influence of cuticle texture on ant ecology.

@@

\begin{table}[h]
  \centering
  \caption{Dataset Subclass Distribution}
  \label{table:classdist_all_full}
  \include{tables/all_samples_full.tex}
\end{table}

\begin{table}[h]
  \centering
  \caption{Dataset Class Distribution}
  \label{table:classdist_rs_full}
  \include{tables/rs_samples_full.tex}
\end{table}

Images were collected and classified by undergraduate students according to the
sculpturing identification protocol. Overall, there are a total of 2443 images
in the dataset. The class distribution with all considered subclasses is shown
in Table \ref{table:classdist_all_full}, and the class distribution with overall
distribution is shown in Table \ref{table:classdist_rs_full}. The relatively
small number of images in the dataset paired with the uneven class distribution
makes it difficult to perform a meaningful classification. However, the goal of
this project is to develop a classification algorithm that can be used to help
classify over 200,000 ant images. Therefore, it is desirable to use a few
representative images from each class to train the algorithm. Additionally, data
augmentation and other methods are performed to increase the number of training
images to create a more robust classification algorithm.
